---
title: "2011 and 2016 Census of Housing and Population Exploratory Analysis"
date: "21 April 2018"
output: html_document
---

```{r setup, include=FALSE}

library(psych)
library(plyr)
library(data.table)
library(ggplot2)

setwd("D:/Education/2015-2018 - Master of Data Science/2018-S1 - Data Science Professional Development 2/AustraliaKidneyDisease")

source("functions.R")

```

## Overview

The following code imports data from the 2011 and 2016 Census of Housing and Population for Australia.

```{r import}

# Import: Indigenous Status by SA2 Level for each state (2011 Census)
sa2Data2011 <- importSA2Data("2011")

# Import: SA2 Level by Indigenous Status, Sex, Age in 10 year blocks, and Weekly Personal Income (2011 Census)
census2011 <- importCensusData("2011", "2011-Census-Analysis/data/2011_Census.csv", "D:/2011_Census.csv")
census2011$Year <- "2011"

# Import: Indigenous Status by SA2 Level for each state (2016 Census)
sa2Data2016 <- importSA2Data("2016")

# Import: SA2 Level by Indigenous Status, Sex, Age in 10 year blocks, and Weekly Personal Income (2016 Census)  
census2016 <- importCensusData("2016", "2016-Census-Analysis/data/2016_Census.csv", "D:/2016_Census.csv")
census2016$Year <- "2016"

# Create new variable for Personal Weekly Income because they are different for 2011 and 2016 Census
uniqueCensus2011 <- unique(census2011$Weekly.Personal.Income)
uniqueCensus2016 <- unique(census2016$Weekly.Personal.Income)
setdiff(uniqueCensus2011, uniqueCensus2016) # Weekly Personal Income Values in 2011 not in 2016
setdiff(uniqueCensus2016, uniqueCensus2011) # Weekly Personal Income Values in 2016 not in 2011
intersect(uniqueCensus2011, uniqueCensus2016) # Shared values

# Not applicable
# Nil Income
# Negative Income
# Not stated
# $1-$299 ($1-$199 & $200-$299 from 2011 data; $1-$149 & $150-$299 from 2016 data)
# $300-$399
# $400-$799 ($400-$599 & $600-$799 from 2011 data; $400-$499, $500-$649 & $650-$799 from 2016 data)
# $800-$999
# $1,000-$1,249
# $1,250-$1,499
# $1,500-$1,999 ($1,500-$1,999 from 2011 data; $1,500-$1,749 & $1,750-$1,999 from 2016 data)
# $2,000 or more ($2,000 or more from 2011 data; $2,000-$2,999 & $3,000 or more from 2016 data)
census2011$Income <- as.character(census2011$Weekly.Personal.Income)
census2011$Income[census2011$Income == "$1-$199" | census2011$Income == "$200-$299"] <- "$1-$299"
census2011$Income[census2011$Income == "$400-$599" | census2011$Income == "$600-$799"] <- "$400-$799"

census2016$Income <- as.character(census2016$Weekly.Personal.Income)
census2016$Income[census2016$Income == "$1-$149" | census2016$Income == "$150-$299"] <- "$1-$299"
census2016$Income[census2016$Income == "$400-$499" | census2016$Income == "$500-$649" | 
                    census2016$Income == "$650-$799"] <- "$400-$799"
census2016$Income[census2016$Income == "$1,500-$1,749" | census2016$Income == "$1,750-$1,999"] <- "$1,500-$1,999"
census2016$Income[census2016$Income == "$2,000-$2,999" | census2016$Income == "$3,000 or more"] <- "$2,000 or more"

census2011$Weekly.Personal.Income <- NULL   
census2016$Weekly.Personal.Income <- NULL   

# Aggregate rows that have been re-coded
attach(census2011)
census2011 <- aggregate(Count ~ . , sum, data=census2011)
detach(census2011)

attach(census2016)
census2016 <- aggregate(Count ~ . , sum, data=census2016)
detach(census2016)

# State population variable
state2011 <- aggregate(Count ~ State, sum, data=census2011)
colnames(state2011)[2] <- "StatePopulation"
census2011 <- merge(census2011, state2011, by="State")  # QLD 2016 4644165

state2016 <- aggregate(Count ~ State, sum, data=census2016)
colnames(state2016)[2] <- "StatePopulation"
census2016 <- merge(census2016, state2016, by="State")

# SA2 population variable
sa22011 <- aggregate(Count ~ SA2, sum, data=census2011)
colnames(sa22011)[2] <- "SA2Population"
census2011 <- merge(census2011, sa22011, by="SA2")

sa22016 <- aggregate(Count ~ SA2, sum, data=census2016)
colnames(sa22016)[2] <- "SA2Population"
census2016 <- merge(census2016, sa22016, by="SA2")

# Merge 2011 and 2016 data into one census dataset
census <- rbind(census2011, census2016)
census$Year <- as.factor(census$Year)
census$Income <- as.factor(census$Income)
#census$Weekly.Personal.Income <- NULL

# Add dummy variables for State, Income, Sex and Age variables
states <- as.data.frame(dummy.code(census$State))
ages <- as.data.frame(dummy.code(census$Age))
incomes <- as.data.frame(dummy.code(census$Income))
census <- cbind(census, states, ages, incomes)
census$Male <- as.numeric(census$Sex == "Male")
#census$Over50Years <-

rm(uniqueCensus2011, uniqueCensus2016, states, ages, incomes, 
   census2011, census2016, sa22011, sa22016, state2011, state2016)

summary(census)   # When transforming the data counts of zero were excluded
describe(census$Count, IQR=TRUE) 

# Create data set containing only the Indigenous population
censusIndigenous <- census[which(census$Indigenous == 1),]
censusIndigenous$Indigenous <- NULL
censusIndigenous$Indigenous.Status <- NULL
attach(censusIndigenous)
censusIndigenous <- aggregate(Count ~ . , sum, data=censusIndigenous)
detach(censusIndigenous)

summary(censusIndigenous)   # When transforming the data counts of zero were excluded
describe(censusIndigenous$Count, IQR=TRUE) 

# Indigenous population data (2011 only)
censusIndigenous2011 <- censusIndigenous[which(censusIndigenous$Year == "2011"),]
censusIndigenous2011$Year <- NULL
censusIndigenous2011$Male <- NULL

# Indigenous population data (2016 only)
censusIndigenous2016 <- censusIndigenous[which(censusIndigenous$Year == "2016"),]
censusIndigenous2016$Year <- NULL
censusIndigenous2016$Male <- NULL

```

### SA2 Level (2011 Census)

```{r sa2_2011}

dim(sa2Data2011) # data frame dimensions
str(sa2Data2011) # structure of data
head(sa2Data2011)

summary(sa2Data2011)
describe(sa2Data2011[c(2:7)], IQR=TRUE) # Summary statistics of population counts in each SA2 area
  
count(sa2Data2011, 'State') # Number of SA2 areas in each state
sum(count(sa2Data2011, 'State')$freq) # Number of SA2 areas in Australia for 2011 Census

```

### 2011 Census Data

```{r census_2011}

census2011 <- census[which(census$Year == "2011"),]
dim(census2011) # data frame dimensions
str(census2011) # structure of data
head(census2011, n=5) 

unique(census2011$State) # Levels of State variables
unique(census2011$Indigenous.Status) # Levels of Indigenous Status variable
unique(census2011$Sex) # Levels of Sex variable
unique(census2011$Age) # Levels of Age variable
unique(census2011$Income) # Levels of Weekly Personal Income variable

summary(census2011)   # When transforming the data counts of zero were excluded
describe(census2011$Count, IQR=TRUE) 
  
sum(census2011$Count) # Number of people counted in 2011 Census
length(unique(census2011$SA2)) # Number of SA2 areas with population counts > 0
length(unique(sa2Data2011$SA2)) - length(unique(census2011$SA2)) # Number of SA2 areas with no population recorded

options(scipen=999) # Disable scientific notation
total <- sum(census2011$Count)
customFunction <- function(x) c(sum = sum(x), percent = (sum(x)/total)*100)

attach(census2011)

# Number of people counted in each state during 2011 Census
aggregate(Count ~ State , FUN = customFunction, data=census2011)

# Number of people by indigenous status
aggregate(Count ~ Indigenous.Status , FUN = customFunction, data=census2011)

# Number of people who are indigenous or not
aggregate(Count ~ Indigenous , FUN = customFunction, data=census2011)

# Number of people by sex
aggregate(Count ~ Sex, FUN = customFunction, data=census2011)

# Number of people by age range
aggregate(Count ~ Age, FUN = customFunction, data=census2011)

# Number of people by Weekly Personal Income
aggregate(Count ~ Income, FUN = customFunction, data=census2011)

# Number of people per state by indigenous status
aggregate(Count ~ Indigenous + State, FUN = customFunction, data=census2011)

# Number of people per Indigenous Status by age range
aggregate(Count ~ Indigenous + Age, FUN = customFunction, data=census2011)

# Number of people per Indigenous Status by Weekly Personal Income
aggregate(Count ~ Indigenous + Income, FUN = customFunction, data=census2011)

detach(census2011)

rm(census2011, total)
options(scipen=0) # Enable scientific notation
  
```

### SA2 Level (2016 Census)

```{r sa2_2016}

dim(sa2Data2016) # data frame dimensions
str(sa2Data2016) # structure of data
head(sa2Data2016)

summary(sa2Data2016)
describe(sa2Data2011[c(2:7)], IQR=TRUE) # Summary statistics of population counts in each SA2 area
  
count(sa2Data2016, 'State') # Number of SA2 areas in each state
sum(count(sa2Data2016, 'State')$freq) # Number of SA2 areas in Australia for 2016 Census

```

### 2016 Census Data

```{r census_2016}

census2016 <- census[which(census$Year == "2016"),]
dim(census2016) # data frame dimensions
str(census2016) # structure of data
head(census2016, n=5) 

unique(census2016$State) # Levels of State variable
unique(census2016$Indigenous.Status) # Levels of Indigenous Status variable
unique(census2016$Sex) # Levels of Sex variable
unique(census2016$Age) # Levels of Age variable
unique(census2016$Income) # Levels of Weekly Personal Income variable

summary(census2016)   # When transforming the data counts of zero were excluded
describe(census2016$Count, IQR=TRUE) 
  
sum(census2016$Count) # Number of people counted in 2016 Census
length(unique(census2016$SA2)) # Number of SA2 areas with population counts > 0
length(unique(sa2Data2016$SA2)) - length(unique(census2016$SA2)) # Number of SA2 areas with no population recorded

options(scipen=999) # Disable scientific notation
total <- sum(census2016$Count)
customFunction <- function(x) c(sum = sum(x), percent = (sum(x)/total)*100)

attach(census2016)

# Number of people counted in each state during 2011 Census
aggregate(Count ~ State , FUN = customFunction, data=census2016)

# Number of people by indigenous status
aggregate(Count ~ Indigenous.Status , FUN = customFunction, data=census2016)

# Number of people who are indigenous or not
aggregate(Count ~ Indigenous , FUN = customFunction, data=census2016)

# Number of people by sex
aggregate(Count ~ Sex, FUN = customFunction, data=census2016)

# Number of people by age range
aggregate(Count ~ Age, FUN = customFunction, data=census2016)

# Number of people by Weekly Personal Income
aggregate(Count ~ Income, FUN = customFunction, data=census2016)

# Number of people per state by indigenous status
aggregate(Count ~ Indigenous + State, FUN = customFunction, data=census2016)

# Number of people per Indigenous Status by age range
aggregate(Count ~ Indigenous + Age, FUN = customFunction, data=census2016)

# Number of people per Indigenous Status by Weekly Personal Income
aggregate(Count ~ Indigenous + Income, FUN = customFunction, data=census2016)

detach(census2016)

rm(census2016, total, customFunction)
options(scipen=0) # Enable scientific notation

```

# Plots of Indigenous Data

```{r Plots}

# 2011 Census Data
plot(sa2Data2011$State, main="Number of SA2 Levels per State", ylab="SA2 Levels")
boxplot(Count ~ Age, data = censusIndigenous2011)
boxplot(Count ~ Income, data = censusIndigenous2011)
pairs(~Age+Income+Sex+Count, data=censusIndigenous2011, main="Simple Scatterplot Matrix (2011)")

# Histogram with number labels
Histogram <- function(data, x, xaxislabel) {
  total <- sum(data$Count)
  attach(data)
  x <- aggregate(Count ~ x , FUN = sum, data=data)
  detach(data)
  
  ggplot(data=x, aes(x=x , y=Count)) + geom_bar(colour="black", stat="identity") + 
  geom_text(aes(label=Count), vjust=2, color="white", fontface = "bold", size=4) + # vjust=1.6
  labs(x=xaxislabel,y="Count") + theme_classic()
}
Histogram(censusIndigenous2011, Income, "Income")
Histogram(censusIndigenous2011, State, "State")
Histogram(censusIndigenous2011, Sex, "Sex")
Histogram(censusIndigenous2011, Age, "Age")

# 2016 Census Data
plot(sa2Data2016$State, main="Number of SA2 Levels per State", ylab="SA2 Levels")
boxplot(Count ~ Age, data = censusIndigenous2016)
boxplot(Count ~ Income, data = censusIndigenous2016)
pairs(~Age+Income+Sex+Count, data=censusIndigenous2016, main="Simple Scatterplot Matrix (2016)")

Histogram(censusIndigenous2016, Income, "Income")
Histogram(censusIndigenous2016, State, "State")
Histogram(censusIndigenous2016, Sex, "Sex")
Histogram(censusIndigenous2016, Age, "Age")

rm(Histogram)

```

# Normality Tests

In this section the normality of count data for the Indigenous population (2011) is examined.

```{r Normality}

# https://www.r-bloggers.com/normality-tests-for-continuous-data/
# http://stat.ethz.ch/R-manual/R-devel/library/stats/html/shapiro.test.html

data <- censusIndigenous2011$Count
sample <- sample(data, size = 5000)
hist(sample)
shapiro.test(sample) # Sample size must be between 3 and 5000

library(nortest)
# The A-D test is susceptible to extreme values, and may not give good results for very large data sets
ad.test(data) # very low p-value indicates that the distribution is not normal

qqnorm(data)
qqline(data, col = "red")

rm(data, sample)


```

# Transformation

For techniques such as correlation analysis and regression analysis a normal distribution is required. Because the population counts are not normal the data needs to be transformed.

If distribution is right-skewed: sqrt(x); log(x); 1/sqrt(x); 1/x

```{r Transformation}

# Transformations: http://rcompanion.org/handbook/I_12.html

library(MASS)
Box = boxcox(censusIndigenous2011$Count ~ 1,              # Transform Turbidity as a single vector
             lambda = seq(-6,6,0.1)      # Try values -6 to 6 by 0.1
             )
Cox = data.frame(Box$x, Box$y)            # Create a data frame with the results
Cox2 = Cox[with(Cox, order(-Cox$Box.y)),] # Order the new data frame by decreasing y
Cox2[1,]                                  # Display the lambda with the greatest
                                          #    log likelihood
lambda = Cox2[1, "Box.x"]                 # Extract that lambda
censusIndigenous2011$TransformedCount = (censusIndigenous2011$Count ^ lambda - 1)/lambda #Transform the original data

#censusIndigenous2011$TransformedCount <- log(censusIndigenous2011$Count)
#censusIndigenous2011$TransformedCount <- sqrt(censusIndigenous2011$Count)
#censusIndigenous2011$TransformedCount <- 1/sqrt(censusIndigenous2011$Count)
#censusIndigenous2011$TransformedCount <- 1/censusIndigenous2011$Count
#censusIndigenous2011$TransformedCount <- sign(censusIndigenous2011$Count) * abs(censusIndigenous2011$Count)^(1/3) 

data <- censusIndigenous2011$TransformedCount
sample <- sample(data, size = 5000)
hist(sample)
shapiro.test(sample) # Sample size must be between 3 and 5000

# The A-D test is susceptible to extreme values, and may not give good results for very large data sets
ad.test(censusIndigenous2011$TransformedCount) # very low p-value indicates that the distribution is not normal

qqnorm(data)
qqline(data, col = "red")

rm(Cox, Cox2, Box, data, lambda, sample)

```

# Correlation

In this section the correlation between variables for the Indigenous population are examined.

```{r Correlation}

# https://stats.stackexchange.com/questions/108007/correlations-with-unordered-categorical-variables
# https://stats.stackexchange.com/questions/119835/correlation-between-a-nominal-iv-and-a-continuous-dv-variable/124618#124618
  
# Correlation for 2011 Census data only
  
  #library(corrgram)
  #corrgram(censusIndigenous2011, order=TRUE, lower.panel=panel.shade,
  #         upper.panel=NULL, text.panel=panel.txt, cor.method = "spearman",
  #         main="Correlation of Numerical Variables")
  
  #corrgram(censusIndigenous2011, order=TRUE, lower.panel=panel.shade,
  #         upper.panel=NULL, text.panel=panel.txt, cor.method = "pearson",
  #         main="Correlation of Numerical Variables")
  
  nums <- sapply(censusIndigenous2011, is.numeric)
  numeric_data <- censusIndigenous2011[ , nums]

  # calculate correlation matrix
  spearman <- cor(numeric_data, use="complete.obs", method="spearman") # can be used when outliers present
  pearson <- cor(numeric_data, use="complete.obs", method="pearson")  # assumes normality
  
  # summarise the correlation matrix
  print(spearman)
  print(pearson)
  
# Correlation for 2011 and 2016 Data

  #nums <- sapply(censusIndigenous, is.numeric)
  #umeric_data <- censusIndigenous[ , nums]
  
  # calculate correlation matrix
  #spearman <- cor(numeric_data, use="complete.obs", method="spearman") # can be used when outliers present
  #pearson <- cor(numeric_data, use="complete.obs", method="pearson")  # assumes normality
  
  # summarise the correlation matrix
  #print(spearman)
  #print(pearson)
  
  rm(nums, numeric_data, spearman, pearson)

```

# Regression

Multiple linear regression is used to get a better understanding of which variables are important in estimating the number of indigenous people.

Assumptions for multiple linear regression:
* Error terms must be Normally distributed
* Linearity: The relationship between y and x is linear
* Independent errors: the residuals are independent
** In particular, repeated observations on the same individual are not allowed
* Normality: the residuals are Normally distributed for any given value of x - use a P-P or Q-Q Plot
* Equal Variance (homoscedasticity): the residuals have constant variance around the 0 line

These assumptions can be checked using plots. If assumptions have been violated, findings cannot be generalised beyond the sample.

A good model has a large F-ratio and a small P-value

In multiple regression, R2 is the square of the multiple correlation coefficient between the dependent variable and the predictors. The adjusted R2 indicates the loss of predictive power or shrinkage

Binary logistic regression: Used to predict a binary (dichotomous) categorical response variable from one or more categorical and/or continuous explanatory variables

```{r Regression}

# https://www.statmethods.net/stats/regression.html
# https://data-flair.training/blogs/r-nonlinear-regression/
# Regression for Count data: http://rcompanion.org/handbook/J_01.html
# Survey package: https://cran.r-project.org/web/packages/survey/survey.pdf

# Stepwise Regression
fit <- lm(Count ~ State + Sex + Age + Income + City + InnerRegional + OuterRegional + Remote + VeryRemote + 
            MigratoryOffshoreShipping + NoUsualAddress, data=censusIndigenous2011)
step <- stepAIC(fit, direction="both")
step$anova # display results

# TODO: Diagnostics
library(car)
Anova(step, type="II")
  
x = residuals(step)
library(rcompanion)
plotNormalHistogram(x)
  
qqnorm(residuals(step), ylab="Sample Quantiles for residuals")
qqline(residuals(step), col="red")

plot(fitted(step), residuals(step))

# All Subsets Regression
#library(leaps)

#attach(censusIndigenous2011)
#leaps<-regsubsets(Count ~ State + Sex + Age + Income + City + InnerRegional + OuterRegional + Remote + VeryRemote + 
#            MigratoryOffshoreShipping + NoUsualAddress, data=censusIndigenous2011, nbest=10)#, really.big=T)
#detach(censusIndigenous2011)

# view results 
#summary(leaps)

# plot a table of models showing variables in each model.
# models are ordered by the selection statistic.
#plot(leaps,scale="r2")

# plot statistic by subset size 
#library(car)
#subsets(leaps, statistic="rsq")

# TODO: Apply to 2016 data

rm(fit, step, leaps, x)

```

# Small Area Estimation

In this section some simple small area estimation models are trained to try and estimate the number of Indigenous people in different SA2 areas. This is improve knowledge in how to train and evaluate the models. More advanced models will be created to estimate the number of Indigenous people with kidney disease using the health survey data.

```{r Estimation}

library(sae) # small area estimation
library(rsae) # robust small area estimation

attach(censusIndigenous2011)
Xs <- cbind(City, InnerRegional, OuterRegional, Remote, VeryRemote) # Construct design matrix for sample elements
domains <- as.numeric(unique(State))
detach(censusIndigenous2011)

attach(censusIndigenous2016)
Xoutsamp <- cbind(as.numeric(State), City, InnerRegional, OuterRegional, Remote, VeryRemote) 
#Xoutsamp <- Xoutsamp[SA2 %in% domains,]
detach(censusIndigenous2016)

#attach(censusIndigenous2011)
#result <- ebBHF(Count ~ Xs, dom=State, Xnonsample=Xoutsamp, 
#                MC=10, constant=3600, indicator=function(y) {y})
#detach(censusIndigenous2011)

# EXAMPLE

  data(incomedata) # Load data set
  attach(incomedata)
  
  # Construct design matrix for sample elements
  Xs <- cbind(age2, age3, age4, age5, nat1, educ1, educ3, labor1, labor2)
  
  # Select the domains to compute EB estimators.
  data(Xoutsamp)
  
  domains <- unique(Xoutsamp[,"domain"])
  
  # Poverty gap indicator
  povertyline <- 0.6*median(income)
  povertyline # 6477.484
  povgap <- function(y)
  {
  z <- 6477.484
  result <- mean((y<z) * (z-y) / z)
  return (result)
  }
  
  # Compute EB predictors of poverty gap. The value constant=3600 is selected
  # to achieve approximately symmetric residuals.
  set.seed(123)
  result <- ebBHF(income ~ Xs, dom=prov, selectdom=domains,
  Xnonsample=Xoutsamp, MC=10, constant=3600, indicator=povgap)
  result$eb
  result$fit$summary
  result$fit$fixed
  result$fit$random[,1]
  result$fit$errorvar
  result$fit$refvar
  result$fit$loglike
  result$fit$residuals[1:10]
  detach(incomedata)
  
  rm(domains, povertyline, result, Xs, Xoutsamp, incomedata, povgap)

```


